\uplevel{%
  \noindent\textbf{Part 2: Multiple Choice}\\
  Please clearly circle your answer.
  If you change your answer, please erase your previous marking cleanly.
  Each choice is graded separately and is worth 1 point.
}


\question[4]
Multi-head attention computes \(H\) attention outputs and then combines them. Which of the following statements is correct? \textbf{Select all that apply.}
\begin{enumerate}
\item Different heads can use different learned projection matrices, which can lead to different attention patterns.
\item Concatenating the head outputs typically increases the feature dimension before a final linear projection maps back to \(\dmodel\).
\item Using more heads removes the need for causal masking in a decoder.
\item Multi-head attention can be  computed in parallel.
\end{enumerate}
\begin{soln}
A,B,D
\end{soln}



% \question[4]
% (\textit{PyTorch-level concept}) A common implementation pattern for attention is:
% \[
% \texttt{scores}\leftarrow \frac{QK^\top}{\sqrt{d_k}},\qquad
% \texttt{scores}\leftarrow \texttt{scores}+\texttt{mask},\qquad
% \texttt{attn}\leftarrow \texttt{softmax}(\texttt{scores}).
% \]
% Which of the following statements is correct? \textbf{Select all that apply.}
% \begin{enumerate}
% \item A causal mask is used in the encoder so that the encoder attends only to previous source tokens. 
% \item Adding \texttt{mask} before softmax is a way to enforce that some \((i,j)\) pairs receive near-zero attention weight.
% \item If \texttt{mask} uses large negative values for disallowed pairs, those pairs receive negligible probability after softmax.
% \item The main purpose of the mask is to change the value vectors \(V\) by setting some entries of \(V\) to zero. 
% \end{enumerate}
% \begin{soln}
% B, D
% \end{soln}



\question[4]
Consider a decoder-only Transformer trained for next-token prediction on sequences \((y_1,\ldots,y_L)\). Which of the following statements is correct? \textbf{Select all that apply.}
\begin{enumerate}
\item During training, we can compute a loss for each position \(t\) using the predicted distribution for \(y_{t+1}\) and the known target token.
\item During inference, cross-attention to an encoder is required for any Transformer model.
\item During inference, tokens are generated sequentially because the next token is unknown until it is produced.
\item During training, causal masking is unnecessary because the targets are known. 
\end{enumerate}
\begin{soln}
A,C
\end{soln}

\question[4]
Consider a decoder--encoder model with decoder states \(S_\dec \in \rR^{L_\dec \times \dmodel}\) and encoder states \(H_\enc \in \rR^{L_\enc \times \dmodel}\). Which statements correctly describe cross-attention? \textbf{Select all that apply.}
\begin{enumerate}
\item Cross-attention typically needs a causal mask to prevent attending to future decoder positions. 
\item The attention weight matrix has size \(L_\dec \times L_\enc\) for each head.
\item For each head \(h\), \(Q^{(h)} = S_\dec W^{Q,h}\) and \(K^{(h)} = H_\enc W^{K,h}\).
\item In cross-attention, both \(K^{(h)}\) and \(V^{(h)}\) are computed from \(H_\enc\).
\end{enumerate}
\begin{soln}
B,C,D

\end{soln}


% \question[4]
% Which of the following statements about training versus inference in a decoder-only Transformer is correct? \textbf{Select all that apply.}
% \begin{enumerate}
% \item In training with teacher forcing, all positions can be computed in parallel within a layer because the target tokens are known.
% \item In inference, generation is sequential because token \(y_{t+1}\) is unknown until the model produces it.
% \item Even in training, causal masking is used so that position \(t\) does not use information from target positions \(>t\) when predicting \(y_{t+1}\).
% \item In inference, causal masking is unnecessary because the model already knows future tokens.
% \end{enumerate}
% \begin{soln}
% A,B,C

% \end{soln}

\question[4]
Consider the following pseudocode for masked attention logits before \(\text{softmax}\):
\[
\text{scores} = QK^\top/\sqrt{d_k},\qquad
\text{scores}[\,\text{mask}=0\,] = -\infty,\qquad
A=\text{softmax}(\text{scores}).
\]
Which statement is correct?
\begin{enumerate}
\item This masking is primarily to make attention computations faster by reducing the asymptotic complexity from \(O(L^2)\) to \(O(L)\).
\item A padding mask is used to increase the influence of PAD tokens so that batch statistics are more stable.
\item Setting masked logits to a very negative value makes the corresponding \(\text{softmax}\) weights close to zero.
\item A causal mask allows a decoder position \(t\) to attend to positions \(>t\) during training.
\end{enumerate}
\begin{soln}
C
\end{soln}

\question[4]
Which of the following statements about Rotary Positional Embedding (RoPE) are correct? \textbf{Select all that apply.}

\begin{enumerate}
\item RoPE is applied only to the value vectors. 
\item In RoPE, attention scores depend on relative positions through $(j-i)$.
\item RoPE changes the norm (magnitude) of query and key vectors.
\item RoPE rotates query and key vectors using position-dependent rotations. 
\end{enumerate}

\begin{soln}
B, D
\end{soln}


\question[4]
Which statements are correct about BERT-style MLM? \textbf{Select all that apply.}
\begin{enumerate}
\item MLM is often applied to decoder-only next-token prediction.
\item Some input tokens are masked and predicted.
\item Both left and right context can be used.
\item MLM is a common pretraining objective for encoder-style models.
\end{enumerate}
\begin{soln}
B, C, D
\end{soln}

\question[4]
If you observe that a fine-tuned model starts repeating phrases, which  statements are correct? \textbf{Select all that apply.}
\begin{enumerate}
\item Keep learning rate fixed and train much longer.
\item Ignore validation behavior and continue training.
\item Lower the learning rate or stop earlier.
\item This observation implies underfitting.
\end{enumerate}
\begin{soln}
C
\end{soln}

% \question[4]
% Which statement correctly describes self-instruct style generation? \textbf{Select all that apply.}
% \begin{enumerate}
% \item It requires full manual authoring of every new instruction.
% \item It is a one-shot process with no iteration.
% \item It never uses seed tasks.
% \item It commonly bootstraps from seed tasks and iteratively adds filtered generations.
% \end{enumerate}
% \begin{soln}
% D
% \end{soln}

\question[4]
Which statements are correct about dataset mixing in instruction tuning? \textbf{Select all that apply.}
\begin{enumerate}
\item Mixture composition can change held-out metrics.
\item Weighting strategies can matter, not just total data size.
\item Equal weights are usually optimal.
\item Mixing has little effect once model size is fixed.
\end{enumerate}
\begin{soln}
A, B
\end{soln}

% \question[4]
% Which habits are appropriate when auditing instruction-tuned model behavior? \textbf{Select all that apply.}
% \begin{enumerate}
% \item Evaluate across held-out tasks or formulations.
% \item Inspect failure cases linked to data quality issues.
% \item Compare results across different data-mixture designs.
% \item Use only training loss to claim broad generalization.
% \end{enumerate}
% \begin{soln}
% A, B, C
% \end{soln}

\question[4]
Which statements align with cautionary interpretation of emergence claims? \textbf{Select all that apply.}
\begin{enumerate}
\item Evaluation design has no effect on perceived thresholds. 
\item A discrete evaluation metric can make gradual improvement appear abrupt.
\item Capability-level emergence and smooth loss scaling can coexist.
\item Capability thresholds on specific tasks can appear even when training loss changes smoothly.
\end{enumerate}
\begin{soln}
B,C,D.
\end{soln}


% multiple choice
\question[4]
Which statements are consistent with scaling-law studies? \textbf{Select all that apply.}
\begin{enumerate}
\item Scaling laws imply that architecture shape is always irrelevant.
\item Scaling results remove the need for empirical evaluation on downstream tasks.
\item Test loss often follows a smooth power-law trend with increased scale.
\item Increasing parameters, data, and compute in a balanced way is usually more effective than scaling only one factor.
\end{enumerate}
\begin{soln}
C,D.
\end{soln}


% \question[4]
% You need to prompt an LLM so that a single concise answer is generated from trusted material
% only. Which setups are most appropriate?
%  \textbf{Select all that apply.}
% \begin{enumerate}
% \item Provide the context text and ask the model to summarize the  relevant information without additional constraints.
% \item Ask for long reasoning traces without final-answer constraint.
% \item We only need to ask the model to answer concisely based on the provided material.
% \item Provide delimited trusted context and require a fixed output format.
% \end{enumerate}
% \begin{soln}
% D.
% \end{soln}


% \question[4]
% Under a fixed compute budget, which strategies are most aligned with compute-optimal scaling observations?
% \textbf{Select all that apply.}
% \begin{enumerate}
% \item Maximize model size while keeping training tokens very small.
% \item Keep model and data near a balanced tradeoff curve rather than overemphasizing one side.
% \item Fully converge the smallest model possible, regardless of data scale.
% \item Use identical token-per-parameter ratios for all model families and settings.
% \end{enumerate}
% \begin{soln}
% B.
% \end{soln}

\question[4]
Which statements about precision choices are accurate?  \textbf{Select all that apply.}
\begin{enumerate}
\item BF16/FP16 are always numerically safer than FP32 for any model and any setup.
\item FP32 is faster than BF16/FP16 on recent tensor-core GPUs.
\item In mixed precision training, BF16/FP16 are the preferred training precision for optimizer states.
\item Precision choices affect both speed and memory footprint.
\end{enumerate}
\begin{soln}
D.
\end{soln}

\question[4]
Which statements about ZeRO-Offload are correct? \textbf{Select all that apply.}
\begin{enumerate}
\item It can lower peak GPU memory by moving optimizer/master states to CPU.
\item It incurs little communication overhead.
\item It may reduce training speed if CPU-GPU bandwidth is the bottleneck.
\item It eliminates the need for mixed precision.
\end{enumerate}
\begin{soln}
A,C.
\end{soln}

\question[4]
Which statements about LoRA design are correct? \textbf{Select all that apply.}
\begin{enumerate}
\item Rank $r$  is a key hyperparameter.
\item LoRA typically uses less optimizer-state memory than full-parameter fine-tuning.
\item LoRA updates all base-model weight entries during fine-tuning. 
\item Applying LoRA only to attention projections is a common practical choice.
\end{enumerate}
\begin{soln}
A,B,D.
\end{soln}

\question[4]
Which statement is most accurate about preference data in RLHF? \textbf{Select all that apply.}
\begin{enumerate}
\item It directly provides gradients for PPO without a reward model.
\item It provides pairwise comparisons that indicate which response is preferred.
\item It is often used  in pretraining.
\item It is the same type of supervision as SFT.
\end{enumerate}
\begin{soln}
B.
\end{soln}

\question[4]
Which claims are consistent with InstructGPT findings? \textbf{Select all that apply.}
\begin{enumerate}
\item Alignment-tuned smaller models can be preferred to much larger untuned baselines.
\item RLHF with pairwise preference can replace SFT.
\item One can mix policy optimization data with pretraining text to better retain broad capabilities.
\item Larger model parameter count alone leads to better human preference ranking.
\end{enumerate}
\begin{soln}
A,C.
\end{soln}

\question[4]
Which statements are true about comparing RLHF and RLAIF ? \textbf{Select all that apply.}
\begin{enumerate}
\item RLHF and RLAIF rely on  the same feedback source at each stage. 
\item Both can include SFT, reward modeling, and policy optimization stages.
\item RLAIF introduces AI feedback to reduce dependence on human comparison data.
\item The objective of maximizing helpfulness does not affect harmlessness.
\end{enumerate}
\begin{soln}
B,C.
\end{soln}

\question[4]
Which statements are correct about RAFT training dynamics? \textbf{Select all that apply.}
\begin{enumerate}
\item Model parameter updates in RAFT require backpropagation through the generation sampling process itself.
\item If one candidate is selected from $K$ samples per prompt, the acceptance ratio is $1/K$.
\item RAFT isn't iterated across multiple stages.
\item Reward-model filtered data are used for supervised fine-tuning updates.
\end{enumerate}
\begin{soln}
B,D.
\end{soln}

% \question[4]
% Which statements correctly compare RAFT and Iterative DPO? \textbf{Select all that apply.}
% \begin{enumerate}
% \item RAFT stage updates are based on selected responses and SFT-style supervision.
% \item Iterative DPO stage updates are based on pairwise preference data.
% \item Both methods can be viewed as iterative policy improvement loops.
% \item Both methods optimize the same pairwise logistic objective at each stage.
% \end{enumerate}
% \begin{soln}
% A,B,C.
% \end{soln}

\question[4]
Which statements are correct about the DPO objective? \textbf{Select all that apply.}
\begin{enumerate}
\item DPO can be written using log-ratios between $\pi_\theta$ and a reference policy $\pi_{\mathrm{ref}}$. 
\item DPO trains a separate reward model. 
\item DPO requires sampling trajectories from an environment with delayed rewards as in standard online RL.
\item DPO training data are pairwise preferences (chosen vs rejected responses) for a prompt. 
\end{enumerate}
\begin{soln}
A,D.
\end{soln}

\question[4]
Which statements are correct about GRPO objective design?
\textbf{Select all that apply.}
\begin{enumerate}
\item GRPO always requires a separately trained value model.
\item GRPO shares normalization statistics across different prompts within the same batch.
\item Advantage construction can come from relative rewards among sampled outputs.
\item KL regularization can be included explicitly in objective design.
\end{enumerate}
\begin{soln}
C,D.
\end{soln}

\question[4]
Which statements are correct about verifier-based rewards in RLVR?
\textbf{Select all that apply.}
\begin{enumerate}
\item They are feasible for any open-ended subjective task without additional design.
\item They still require careful checker/reward design to avoid loopholes.
\item They can be combined with a small format/structure reward when needed.
\item They are generally suitable for problems with clear correctness criteria.
\end{enumerate}
\begin{soln}
B,C,D.
\end{soln}

% \question[4]
% Which statements are correct about stage-wise reasoning-model training?
% \textbf{Select all that apply.}
% \begin{enumerate}
% \item A staged pipeline can alternate objectives that prioritize different qualities.
% \item Using only one static supervised stage is always sufficient for best reasoning and alignment.
% \item Reward quality and prompt/data mixture can materially affect outcomes.
% \item Later alignment-focused stages can coexist with preserved reasoning strength.
% \end{enumerate}
% \begin{soln}
% A,C,D.
% \end{soln}


% Local Variables:
% TeX-master: "../exam.tex"
% End:

%  LocalWords:  pseudocode logits QK softmax RoPE MLM pretraining PPO
%  LocalWords:  underfitting GPUs ZeRO LoRA RLHF SFT InstructGPT DPO
%  LocalWords:  untuned RLAIF backpropagation GRPO verifier RLVR
