\uplevel{%
  \noindent\textbf{Part 1: True/False}\\
  Please clearly circle your answer. Each question is worth 2 points.
  If you change your answer, please erase your previous marking cleanly.
}





\question[2]
In cross-attention in an encoder--decoder Transformer, queries come from encoder states, while keys and values come from decoder states.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
B
\end{soln}



\question[2]
Encoder-only pretraining is mainly aimed at learning contextual representations rather than left-to-right generation.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
A
\end{soln}

\question[2]
If training loss decreases during fine-tuning, performance on unseen prompts  will improve.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
B
\end{soln}


% true/false
\question[2]
  Increasing task diversity in instruction tuning data generally improves performance on unseen tasks, but may hurt performance if data quality is low or distribution is mismatched.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
A
\end{soln}


\question[2]
If a downstream metric shows threshold-like behavior, that usually implies a discontinuity in training loss.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
B
\end{soln}

\question[2]
Optimizer states are typically more memory-intensive than model parameters when using Adam-style optimizers.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
A
\end{soln}

\question[2]
A reward model in RLHF is often trained with a logistic objective on chosen vs. rejected responses.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
A
\end{soln}

\question[2]
  In KL-regularized policy learning, the reference policy term balances reward maximization and diversity by softly encouraging deviation from the reference model.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
B.
\end{soln}

\question[2]
In DPO, policy updates use chosen-versus-rejected preference pairs as well as a reward model.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
B.
\end{soln}

\question[2]
In RLVR with a reliable verifier,  one can often avoid learning an additional reward model.
\begin{enumerate}
\item True
\item False
\end{enumerate}
\begin{soln}
A.
\end{soln}



% Local Variables:
% TeX-master: "../exam.tex"
% End: