\textbf{Instructions.}
\begin{itemize}
\item
   Homework is due \textbf{\duedate};
   a late penalty of $5\%$ of the maximum score per day will be applied to submissions made after the deadline. The penalty is applied at \textbf{midnight} each day and accumulates daily. After \lastduedate, submissions will no longer be accepted, and no credit will be awarded for the homework. Only your \textbf{most recent submission} will be considered for grading.

\item
  Everyone must individually submit the code component on Gradescope under \hw{} and the written component (PDF) on Canvas.

\item
  The written submission for \hw{} \textbf{must be typed}. You may use \LaTeX, markdown, Google Docs, or any other editor that produces a legible exported file.

\item
  Please make sure your NetID is clear on the first page of the homework.

\item
  Your solution \textbf{must} be written in your own words. You may consult external references, but you must cite them, and your final writeup must still be your own.

\item
  We reserve the right to reduce the auto-graded score for \hwc{} if we detect attempts to game the autograder instead of implementing the intended methods.

\item
  There are \textbf{no regrade requests} on \hwc{}, but you may resubmit as many times as you like before the deadline. Start early and report issues on Campuswire.

\item
  Submit to Gradescope \texttt{\hwc{}}:
  \begin{itemize}
  \item \texttt{gan.py}
  \item \texttt{diffusion.py}
  \item \texttt{outputs/gan\_model.pt}
  \item \texttt{outputs/diffusion\_model.pt}
  \end{itemize}
  If the template asks for a reward or training trace file, submit it using the exact starter-code filename.

\item
  Methods and functions in the template and utility code include docstrings describing the inputs and outputs. The autograder relies on correct implementations of these methods. Follow the docstrings closely.

\item
  Students are permitted to use LLMs, such as ChatGPT or Copilot, to help with the homework. If you do, disclose where and how they were used, and cite any external resources you relied on.

\item
  Please check Campuswire regularly for homework clarifications.
\end{itemize}
