
\textbf{Instructions.}
\begin{itemize}
\item
   Homework is due \textbf{\duedate}; 
   A late penalty of $5\%$ of the maximum score per day will be applied to submissions made after the deadline. The penalty is applied at \textbf{midnight} each day and accumulates daily (e.g., a submission 2 days late will incur a $10\%$ penalty, 3 days late a $15\%$ penalty, and so on). After \lastduedate, submissions will no longer be accepted, and no credit will be awarded for the homework. Note that only your \textbf{most recent submission} will be considered for grading. If your first submission is on time but a later submission is made after \duedate, the late penalty will be applied based on the timestamp of the final submission.


    \item
    Everyone must individually submit the code component on Gradescope under \hw{} and the written component (PDF) on Canvas.

    \item
    The ``written'' submission at  \hw{} \textbf{must be typed}, and submitted in
    any format canvas accepts (to be safe, submit a PDF).  You may use \LaTeX, markdown,
    google docs, MS word, whatever you like; but it must be typed!
    \textbf{We reserve the right to deduct points for handwritten or otherwise illegible submissions.}

    \item
    Please make sure your NetID is clear and large on the first page of the homework.

  \item
    Your solution \textbf{must} be written in your own words. You are expected
    to work on the problems independently. You may use external
    references, but any such reference must be properly cited, and the
    final solution must be written in your own words.
    
    \item
    We reserve the right to reduce the auto-graded score for \hwc{}
    if we detect funny business (e.g., rather than implementing an algorithm,
    you keep re-submitting the assignment to the auto-grader, eventually completing
    a binary search for the answers).

    \item
    There are \textbf{no regrade requests} on \hwc{} which is the code auto-grader;
    however, you can re-submit and re-grade as many times as you like before the deadline!
    Start early and report any issues on Campuswire.

    \item 
    You are required to submit \texttt{gru.py} and \texttt{vae.py}, along with the saved model weights, to the Gradescope \texttt{\hwc{}} assignment. Please ensure that the filenames of the saved model weights are not altered, as changing them may result in loss of points for that question.
    Do not submit any other files.

    \item
    Methods and functions in the template and utility code include
    docstrings describing the inputs and outputs.  The autograder
    relies on correct implementations of these methods.  Follow the
    docstrings to avoid failing tests.

    \item 
    Students are \textbf{permitted} to use LLMs, such as ChatGPT or Copilot, to streamline
    the completion of homework. However, students are required to disclose where and how LLMs
    were used, as well as provide references for any external resources used. 

    \item Please check Campuswire regularly for updates regarding the homework.
\end{itemize}
