\begin{Q}
\textbf{Supervised Fine-Tuning (SFT) Pipeline with LLM-as-Judge. 50 Points}

In this problem, implement a two-part SFT pipeline:
\begin{itemize}
\item Part 1: generate instruction-tuning data with a simplified
  self-instruct loop. Save to \texttt{data/sft\_dataset.json}.
\item Part 2: train gpt2 on \texttt{data/sft\_dataset.json} with
  prompt masking and evaluate with an LLM judge on
  \texttt{data/eval\_dataset.json} (which is provided).
\end{itemize}

You should fill in \emph{only} functions whose docstring begins with \textbf{Implement}.  
Required functions:
\[
\texttt{self\_instruct},\ \texttt{generate\_data},\ \texttt{evaluate\_responses},\ \texttt{tokenize\_with\_prompt\_masking}.
\]

\medskip
\noindent\textbf{Important notes.}
\begin{itemize}
\item Keep all function signatures unchanged.
\item Follow the docstring requirements exactly.
\item The code will use \texttt{gpt-5-nano} for data generation and
  LLM as judge. So you should put your openAI API key into
  \verb|~/.env| for
  the program to load. The overall API cost should be well below \$1 for the problem. 
\end{itemize}

\begin{enumerate}

\item \textbf{Part 1: self-instruct data generation (20 points).}

\begin{enumerate}
\item \textbf{\texttt{self\_instruct(...)} (10 points).}
Implement one simplified self-instruct iteration:
\begin{itemize}
\item sample 8 existing examples,
\item ask the provided LLM for 8 new examples in JSON,
\item keep only generated items with rating $\ge 4$,
\item append accepted items back into the dataset list.
\end{itemize}
Your implementation should be robust to JSON parsing failures (do not crash).

\item \textbf{\texttt{generate\_data(...)} (10 points).}
Generate a dataset of exactly \texttt{num} rows in the format
\[
\{\texttt{"prompt"}:\ldots,\ \texttt{"response"}:\ldots\}.
\]
Convert each accepted self-instruct example as follows:
\begin{itemize}
\item if \texttt{input==""}, use \texttt{prompt = instruction},
\item otherwise use \texttt{prompt = instruction + "\\textbackslash nInput: ..."},
\item use \texttt{response = output}.
\end{itemize}
Return a Hugging Face \texttt{Dataset} with exactly \texttt{num} examples.
\end{enumerate}

\item \textbf{Part 2: SFT with prompt masking and judge evaluation (20 points).}

\begin{enumerate}
\item \textbf{\texttt{tokenize\_with\_prompt\_masking(...)} (10 points).}
Format each training sample as a conversation
(\texttt{user(prompt)} $\rightarrow$ \texttt{assistant(response)}), then construct labels so that:
\begin{itemize}
\item use the provided local conversation formatter,
\item non-assistant tokens (prompt/template/role tokens) are masked with \texttt{-100},
\item assistant response tokens remain unmasked,
\item returned features include \texttt{input\_ids}, \texttt{attention\_mask}, and \texttt{labels}.
\end{itemize}
This enforces instruction tuning loss only on reference responses.

\item \textbf{\texttt{evaluate\_responses(...)} (10 points).}
Implement pairwise evaluation with order-bias reduction:
for each example, evaluate both orderings
\texttt{(field\_1, field\_2)} and \texttt{(field\_2, field\_1)},
then average as in the docstring to produce a win rate in $[0,1]$ for \texttt{field\_1}.
\end{enumerate}

\item \textbf{Run experiments and report (10 points).}

Run your pipeline and report:
\begin{enumerate}
\item Submit program outputs for both Part I and Part II, and the sft
  data 
  \texttt{data/sft\_dataset.json} you generated.
\item Part II evaluation accurac: report LLM judged win rate of SFT
  over base model on the eval set as a percentage (e.g.,
  \texttt{63.5\%}).
\end{enumerate}

\begin{solution}

Program outputs under \texttt{code/reference\_output\_sft.txt}.

Sft data in \texttt{code/sft/data/sft\_dataset.json}.

Win rate (SFT over base): 0.8239.

Students should achieve at least 0.75 to get full score. 

\end{solution}

\end{enumerate}
\end{Q}

%%% Local Variables:
%%% TeX-master: "../hw.tex"
%%% End:
