\begin{Q}
\textbf{RLHF: Reward Modeling and DPO. 50 Points}

In this problem, you will implement two RLHF reward models on pairwise preference data:
\begin{itemize}
\item Bradley--Terry reward model by fineuting a pretrained LLM with
  classification head. 
\item DPO implied reward model with fine tuned LLM over a fixed reference LLM.
\end{itemize}

You should fill in \emph{only} functions whose docstring begins with \textbf{Implement}.  
Required methods:
\[
\texttt{BTTrainer.compute\_loss},\ \texttt{BTTrainer.prediction\_step},
\]
\[
\texttt{DPOTrainer.compute\_loss},\ \texttt{DPOTrainer.prediction\_step}.
\]

\medskip
\noindent\textbf{Dataset and setup.}
Use training data from \texttt{data/train\_prefs.txt} and test data \texttt{data/test\_prefs.txt}.
 
\medskip
\noindent\textbf{Important notes.}
\begin{itemize}
\item Keep all function signatures unchanged.
\item Follow the docstring requirements exactly.
\end{itemize}

\begin{enumerate}

\item \textbf{Part 1: Reward model (Bradley--Terry) (25 points).}

\begin{enumerate}
\item \textbf{\texttt{BTTrainer.compute\_loss(...)} (10 points).}
Given a batch containing chosen/rejected tokenized responses:
\begin{itemize}
\item score chosen and rejected with the reward model,
\item compute difference $d = r_{\text{chosen}} - r_{\text{rejected}}$,
\item compute Bradley--Terry loss using BCE-with-logits target 1:
\[
\mathcal{L}_{\mathrm{BT}} = \mathrm{BCEWithLogits}(d, 1).
\]
\end{itemize}
Return loss, and when requested also return outputs consistent with \texttt{Trainer} expectations.

\item \textbf{\texttt{BTTrainer.prediction\_step(...)} (10 points).}
Implement evaluation-time prediction logic:
zq\begin{itemize}
\item run under \texttt{torch.no\_grad()},
\item return \texttt{(loss, predictions, None)},
\item when \texttt{prediction\_loss\_only=True}, return only loss.
\end{itemize}

\item \textbf{Train and report reward-model evaluatoin metrics (5
    points).}
Submit program outputs.   
Report train/test loss and accuracy before and after training below. 

\begin{solution}

  train\_before: loss=1.086595, acc=0.479800

  test\_before: loss=1.098495, acc=0.487750

  train\_after: loss=0.355826, acc=0.823000

  test\_after: loss=0.732526, acc=0.619500
  
  Students should report test\_after accuracy at least 0.58 to get full score.
\end{solution}

\end{enumerate}

\item \textbf{Part 2: DPO trainer (25 points).}

\begin{enumerate}
\item \textbf{\texttt{DPOTrainer.compute\_loss(...)} (10 points).}
Use the provided \texttt{DPOReward} module to compute implicit rewards for chosen and rejected responses,
then optimize with pairwise BCE-with-logits on reward differences.

\item \textbf{\texttt{DPOTrainer.prediction\_step(...)} (10 points).}
Implement evaluation-time prediction behavior consistent with Hugging Face \texttt{Trainer}.

\item \textbf{Train and report DPO metrics (5 points).}
Submit program outputs.   
Report train/test loss and accuracy before and after training below. 
 \end{enumerate}

\begin{solution}

  train\_before: loss=0.693147, acc=0.500000

  test\_before: loss=0.693147, acc=0.500000

  train\_after: loss=0.594382, acc=0.687800

  test\_after: loss=0.662867, acc=0.592000

  Students should report test\_after accuracy at least 0.58 to get full score.

\end{solution}

\end{enumerate}
\end{Q}

%%% Local Variables:
%%% TeX-master: "../hw.tex"
%%% End:
